---
title: "Benoit_Assignment_3"
author: "Gabrielle"
date: '2022-11-09'
output: 
  html_document:
    code_folding: hide
---

##### Covers lab 4, 5, and 6 and the R markdown formatting discussed in Lecture 12

- why you chose to do the test you did? did you inspect the data and did it meet the assumptions?
- there are multiple answers for all of these... you could do a transformation, but you do not need to

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE,results=FALSE)

library(tidyverse)
library(here)
library(ggpubr) # for some graphic applications that extend ggplot2
library(janitor)
#install.packages("lsr")
library(lsr)
library(effsize)
```

## Question 1:
In order to determine if the current population in the San Antonio River in Santa Barbara County is different from the historical record, I did a one sample t-test, and set mu to equal 4.2 cm (rather than the default, mu = 0), the value from 1955 document that reported the population mean length of adult tidewater gobies. I did this because I am testing to see if there is a statistical difference between the historical mean length of gobies, and the current mean length of gobies (3.5 cm). The one-sample t-test allows us to compare the mean from sample data to a value, usually testing literature point estimates. This test does assume equal variance. 
We found a statistical difference between the current San Antonio Goby fish mean length, and the historical 1955 Goby population mean length (t(9 df), p = 0.0008873), ⍺ = 0.05.
CI: 3.189284, 3.830716

```{r, message=FALSE}
# San Antonio river gobies in centimeters  
# data input: creating a vector of data
san_ant_gobies <- c(2.7, 3.2, 3.2, 3.3, 3.5, 3.6, 3.6, 3.7, 4.0, 4.3)

summary(san_ant_gobies)

san_ant_gobies_one_t <- t.test(san_ant_gobies, mu=4.2)
san_ant_gobies_one_t
```


## Question 2:
Are the tidewater goby lengths in question 1 normally distributed? Provide a figure with caption **or** critical statistical test and justify your results and conclusion.

For this data, to look at whether the tidewater goby lengths are normally distributed, we're going to look at a qq plot. It should be mostly a straight line, however, due to there being very few observations,  some wobble to the line is expected. 

```{r goby_1, fig.align='center',fig.cap="Figure 1. Quantile-Quantile plot showing the empirically observed quantiles of San Antonio River Gobies (cm) as a function of quantiles expected from a normal distribution with the same mean and variance as the empirical distribution."}
# qqplot for san antonio river gobies
qqplot_san_ant_gobies <- ggplot(data.frame(san_ant_gobies), 
                     aes(sample = san_ant_gobies)) +
                     geom_qq() +
                     labs(title = "San Antonio River Gobies") +
                     theme_minimal() + 
                     stat_qq() +
                     stat_qq_line() 

qqplot_san_ant_gobies
```


## Question 3:
The UC Tobacco Free Campus Initiative wants to know if the UC reserves have more or less cigarette butt load as urban open spaces. 
Use your knowledge of statistics to:
1. justify the appropriate statistical test to use
2. explain the results and conclusion of the test you selected
3. make a figure with caption to best explain the data and results.

parametric or non parametric? t test or anova? do preliminary investigations?
do not throw ten plots... say I looked at the residuals, looked at the ... then, I decided
my p value was this, my alpha was this... 
in this figure... 

```{r}
uc_reserves <- c(4, 1, 8, 0, 0, 0, 0, 23, 28, 0, 0, 0, 4, 0, 3, 0, 0, 2, 0, 5, 0, 0, 0, 0, 0)

urban_open <- c(33, 9, 9, 7, 2, 38, 9, 20, 0, 39, 18, 41, 25, 164, 37, 60, 55, 45, 14, 3, 69, 60, 13, 1)

summary(uc_reserves)

summary(urban_open)
```

```{r}
qqplot_uc_reserves <- ggplot(data.frame(uc_reserves), aes(sample = uc_reserves)) +
  geom_qq() +
  labs(title = "UC Reserves") +
  theme_minimal()

qqplot_uc_reserves

qqplot_urban_open <- ggplot(data.frame(urban_open), aes(sample = urban_open)) +
  geom_qq() +
  labs(title = "Urban Open Space") +
  theme_minimal()

qqplot_urban_open

```


```{r}
boxplot_uc_reserves <- ggplot(data = data.frame(uc_reserves), aes(x = uc_reserves)) +
  geom_boxplot() +
  theme_minimal()

boxplot_uc_reserves
# ^ looks very strange

density_overlay_uc_reserves <- ggplot(data = data.frame(uc_reserves), aes(x = uc_reserves)) +
  geom_histogram(aes(y = stat(density))) +
  geom_density(col = "red", size = 2)

density_overlay_uc_reserves
# ^ this is showing me that the data is skewed

t.test(uc_reserves, urban_open)

cohensD(uc_reserves, urban_open)
cohen.d(uc_reserves, urban_open)

```
Our p-value = 0.0005326. It is <0.05, our alpha level of significance, so we can reject our null hypothesis and conclude that there is a statistical difference between the mean of the two groups, for this question, UC Reserves and Urban Open Space. 
Cohen's d shows the effect size, which can tell us how large of a difference this actually is. For this question, a cohen's d of 1.16 communicates that: the average cigarette butt load of the UC reserves is 1.16 standard deviations lower than the average cigarette butt load in urban open spaces, which is considered a large effect size. 




## Question 4:
four groups: 15C, 20C, 25C, 30C
is there a different in how long it takes for them to hatch depending on the temperature
how did you look at the data?
what did you find?
provide a nice figure?
```{r}

```


